\textls[-15]{\thetitle \mbox{(Technical Report)}}

### 2022-07-16 syntax.tex
Warmup section rewritten Different topics for Warmup is stratified for accessibility make first four judgements into warmup section

We highlight some specialities of declarative style in QIIT that differ from a
conventional setting: 
(1) This declarative formulation of type theory does not
have an operational semantics but only judgemental equality (e.g. $\beta$-reduction is now $\beta$-equality).\footnote{Later we will prove
canonicity which can imply a certain level of computational ability of
our theory} 
(2) We directly use the equality from QIIT to represent judgemental equality in our syntax, relieve from us the burden to explicit mentioning equivalence closure; however, this also mean that, \textit{every piece of the syntax data}, including all kinds of judgements (QIIT type) and their derivations will have equality information upon them  
%We will recover canonicity afterwards, which implies the computational ability of our theory. 
% (3) We are dealing with quotient data, so we are actually manipulating
% \textit{equivalence classes of terms} instead of one representative (term) from the equivalence class.
(3) Instead of using meta-level substitution, we use explicit
substitution and de Bruijn indices in our intrinsic formulation. This is
also called \textit{substitution calculus} in the literature and is favoured due
to its closedness to categorical semantic. We will go through these details one-by-one.

@@

% omitting presupposition
Due to our intrinsic setting, \textbf{we can and will omit a lot of premises of the typing rules}~(the presuppositions). 
For example, we don't explicitly
spell out that the rule \ruleref{Type Universe} (that is stating a universe is a proper type) requires a well-formed
context, even though it does implicitly inferrable, \textbf{because it is not
possible to have ill-formed(ill-typed) contexts in our intrinsic setting}. This is
a bit unusual for conventional extrinsic setting where they will have
well-formedness of context everywhere in most of the term judgements. We
consider this as one advantage of using intrinsic setting---this
``well-formed context'' condition is usually attached to some other premises thus
inferrable and can be considered as ``boilerplate code''. Free of
such boilerplate code, the intrinsic setting provides a more concise
representation.

@@

% Example : how judgemental equality is used for the universe
We review the concept of ``Universe'' and at the same time showing how equality is representing semantic. In dependent type setting, function now can return type, of which the type is ``type of type'', i.e., \textit{the universe} $\cU$. We follow the style of \textit{type universe à la Tarski}~\cite{hofmann1997syntax} and
distinguish types from their \textit{codes or names} as showin in \ruleref{Code/Get Type}: given a \textit{code} (term) $\goodTerm{\Gamma}{T}{\cU}$ we can retrieve the corresponding type
$\goodType{\Gamma}{El~T}{}$, and vice-versa.

In other words, (1) a universe $\cU$ is a type including the \textit{codes or names} of types.
(2) a code (term) $\goodTerm{\Gamma}{T}{\cU}$ is not allowed to locate at type position
(i.e., at the rightmost position of a term judgement) but $El \ T$ is (i.e. we distinguish the code and the type). (3)
"c" is the ``inverse'' of $El$, obtaining  the \underline{"c"}ode of a
type, with the following \textit{two (omitted) equations representing their semantic} $"c" (El~T) \equiv T$ and $El
("c"~T) \equiv T$---notice how the two sides of each equality is term of the same type thus a well-typed equality. More concretely,

@@

\underline{\textit{\textbf{Substitution Calculus.}}}
Now we review the substitution calculus.
Substitution is very (if not the most) important because (dependent) functions are
first-class citizens, and lambdas can wrap (almost) every parts and can return
(almost) everything---once we apply the functions,  \textbf{the substitution will happen and needs to
commute} with most all type/term constructs in the function body.

A term $\goodTerm{\Gamma}{t}{A}$ can be understood as one term $t$ of
type $A$, with a bunch of holes (of type $\Gamma$) inside the term $t$.
A substitution $\goodSub{\Gamma}{\gamma}{\Delta}$ can be similarly
understood in the way that $\gamma$ will output a \textbf{series} of
(dependently-typed related) terms of types $\Delta$, parametrized by
$\Gamma$ (i.e., we can consider there are holes of type $\Gamma$ in $\gamma$).
Thus (1) \ruleref{Sub Extend} looks like concatenating one
series of terms with a new term (the weird looking type of
$\goodTerm{\Gamma}{t}{A[\sigma]}$ in the premise is due to our need of type $A$ in
the context $\Delta$ so that we can use \ruleref{Context Extension} to
create new context); (2) we have \ruleref{Proj Subst-1} to extract a
subpart of a given substitution; (3) \ruleref{Proj-Ext} explains why
substitution really looks like a (deeply-nested) pair.

@@

\subsubsection{\textbf{Core Syntax of the Extended MLTT}}
\begin{figure}[!htb]
  \begin{align*}
&\sigma~:~"Sig"~\Gamma~n&&\mathboxtext{\goodSig{\Gamma}{\sigma}{n}}  &&\tau~:~"WSig"~\Gamma~n&&\mathboxtext{\goodWSig{\Gamma}{\tau}{n}} \\ &h~:~"Inh"~\Gamma~\sigma_1~\sigma_2&&\mathboxtext{\goodInh{\Gamma}{h}{\sigma}{\tau}} && f~:~"Seal"~\Gamma~\sigma~T&&\mathboxtext{\goodSeal{\Gamma}{f}{\sigma}{A}}\YZ{A little unfriendly to the reader that there are three different things that sigma can range over. Besides, later you also use gamma to range over substitutions and s to range over W-types, which add to the confusion.}\EDJreply{Fair point. I have fixed fix sigma and tau to be about signatures and W signature, and use gamma as substitutions}
  \end{align*}
\caption{Judgement forms, as Agda types, continued}
\end{figure}
\begin{figure}[!htb]
  \begin{minipage}[b]{0.3\linewidth}
      $$
      \Rule[name=Sig]
      {\goodCtx{\Gamma}{i} \quad n \in \nat}
      {\goodSig{\Gamma}{\_}{n} \text{ is well-formed}}
      $$
      $$
      \Rule[name=Seal]
      {\goodCtx{\Gamma}{i} \quad \goodSig{\Gamma}{\sigma}{n} 
      \quad \goodType{\Gamma}{A}{} }
      {\goodSeal{\Gamma}{\_}{\sigma}{A} \text{ is well-formed}}
      $$\YZ{Doesn't the left column show well-formedness rules for ONLY TWO of the four new judgments? Besides, are these rules used anywhere in the following? Seems a bit unusual to have well-formedness rules for judgments per se.}\EDJreply{Yes. Right column show all the well-formedness rule for all the judgement. That is unusual because it is a direct translation from QIIT. }
  \end{minipage}
  \begin{minipage}[b]{0.6\linewidth}
    \begin{minted}[]{agda}
      data Sig  : Con → ℕ → Set
      data WSig : Con → ℕ → Set 
      data Inh  : (Γ : Con) → Sig Γ n → Sig Γ m → Set
      Seal : (Γ : Con) → Sig Γ n → Sig Γ n → Set
    \end{minted}
  \end{minipage}

@@

The core language builds on MLTT and adds four new judgement forms.
Among them, sealing judgement $\goodSeal{\Gamma}{\_}{\sigma}{A}$ is just a
shorthand for term judgement
$\goodTerm{\Gamma , \cC \sigma}{\_}{A[\pi_1]}$ where $A$ is a type.

@@

We will have dependent function types
and dependent pair types. Dependent pair type is used to model Coq
module, which is useful when we need to model the ``compilation'' from
linkages to modules. We also have $\beta,\eta$-rule for both dependent
function type and dependent pair type.
The substitution law for these two
require \textit{de Bruijn lifting}: for a given substitution
$\goodTerm{\Gamma}{\sigma}{\Delta}$ and a type
$\goodType{\Delta}{A}{}$, we define its de Bruijn lifting
$\goodTerm{\Gamma,A[\sigma]}{\sigma^\uparrow := (\sigma \circ \pi_1, \pi_2)}{\Delta, A}$ . Intuitively speaking, de Bruijn lifting is to ``fill all the holes except the last one''.

$$
\quad 
\Rule[name=Func/DPair Subst]
{\goodSub{\Gamma}{\gamma}{\Delta}
\quad \goodType{\Delta}{A}{} 
\quad \goodType{\Delta, A}{B}{}
}
{
  \goodType{\Gamma}{(\Pi A B)[\gamma] \equiv \Pi A[\gamma] B[\gamma^\uparrow] }{j \cup k}
  \quad 
  \goodType{\Gamma}{(\Sigma A B)[\gamma] \equiv \Sigma A[\gamma] B[\gamma^\uparrow] }{j \cup k}
}
$$
%\YZ{What is the requirement on Delta?}
%\EDJreply{ Delta here is of type Con and thus Delta here should be a well-typed Context. This is another issue I want to talk about, in Agda these kinds of stuff is inferred (so the reader can infer them as well). For example, we know delta is at the place of context, so it is a context, but we don't need to specify that it is a ``well-typed context'' because every context is well-typed in this intrinsic-typing. What's more, I think adding this extra judgement here just cause significant code bloat (like what I complained to you before I know how to use implcit variable in agda. For example, TYPE SUBST, there will be two more judgements (with no more information there) ) (Another reason I prefer (pseudo) agda formulation) }\YZreply{I was thinking about 'Delta |- A'. I suppose it is also implicitly enforced via intrinsically typed syntax. Seems to me it is worth stating, however, or otherwise the rule appears to say that Delta can be any context as long as Gamma |- gamma : Delta.}\EDJreply{You are absolutely right! I think Delta |- A, even though inferrable, mentioning it is good, also some information are B is also ok, please check, resolve if ok, thanks!}
$$
\Rule[name=Type Subst]
{\goodType{\Delta}{T}{j} 
  \quad {\goodSub{\Gamma}{\gamma}{\Delta}}}
{\goodType{\Gamma}{T[\gamma]}{j}}
\quad
\Rule[name=Base Type Subst]
{\goodSub{\Gamma}{\gamma}{\Delta}}
{\goodType{\Gamma}{\cU[\gamma] \equiv \cU }{j + 1} \quad
  \goodType{\Gamma}{\cB[\gamma] \equiv \cB}{0} \quad 
  \goodType{\Gamma}{\bot[\gamma] \equiv \bot}{0}
}
$$
\judgebox{ \goodTerm{\Gamma}{t}{T} }
$$
\Rule[name=Code/Get Type]
{\goodType{\Gamma}{T}{j}}
{\goodTerm{\Gamma}{"c" \ T}{\cU}
}\quad
\Rule[]
{\goodTerm{\Gamma}{T}{\cU}}
{\goodType{\Gamma}{El \ T}{j}}
\quad
\Rule
{}
{\goodTerm{\Gamma}{"tt", "ff"}{\cB}}
\quad 
\Rule[name=Term Subst]
{\goodTerm{\Delta}{t}{T}
  \quad {\goodSub{\Gamma}{\gamma}{\Delta}}}
{\goodTerm{\Gamma}{t[\gamma]}{T[\gamma]}}
$$
$$
\Rule[]
{\goodTerm{\Gamma, A}{t}{B}}
{\goodTerm{\Gamma}{\lambda t}{\Pi A B}}
\quad 
\Rule[]
{\goodTerm{\Gamma}{u}{A} 
\quad \goodTerm{\Gamma}{v}{B[(id, u)]}}
{\goodTerm{\Gamma}{(u,v)}{\Sigma A B}}
\quad 
\Rule[]
{\goodTerm{\Gamma}{t}{\Sigma A B}}
{\goodTerm{\Gamma}{"pjl" \ t}{A}
\quad \goodTerm{\Gamma}{"pjr" \  t}{B[(id, "pjl" \  t)]}
}
$$
$$
\Rule
{}
{\goodTerm{\Gamma}{(\lambda t)[\gamma] \equiv \lambda (t[\gamma^\uparrow])}{\Pi A B}
\quad \goodTerm{\Gamma}{(u,v)[\gamma] \equiv (u[\gamma],v[\gamma])}{\Sigma A B}
\quad \goodType{\Gamma}{El \ (T[\gamma]) \equiv (El \ T) [\gamma]}{}
}
$$

$$
\Rule[name=Base Type/Term Subst]
{\goodSub{\Gamma}{\gamma}{\Delta}}
{\goodType{\Gamma}{(El \ T)[\gamma] \equiv El \ (T[\gamma]) }{j} \quad
 \goodTerm{\Gamma}{"tt"[\gamma] \equiv "tt"}{\cB} \quad 
 \goodTerm{\Gamma}{"ff"[\gamma] \equiv "ff"}{\cB} 
}
$$
\judgebox{\goodSub{\Gamma}{\sigma}{\Delta}}
$$
\Rule[name=Ept Subst]
{}{\goodSub{\Gamma}{\epsilon}{\cdot}}
\quad
\Rule[]
{}{\goodSub{\Gamma}{"id"}{\Gamma}}
\quad
\Rule[name=Sub Comp]{
  \goodSub{\Delta}{\delta}{\Theta}
  \quad \goodSub{\Gamma}{\gamma}{\Delta} 
}{\goodSub{\Gamma}{\delta \circ \gamma}{\Theta}}
\Rule[name=Sub Extend]
{\goodSub{\Gamma}{\sigma}{\Delta} \quad \goodTerm{\Gamma}{t}{A[\sigma]}}
{\goodSub{\Gamma}{(\sigma, t)}{(\Delta, A)}}
$$

We have shown some exemplar rules for all four standard types of
judgments. Now we focus on the newly introduced facility. We have two kinds of signatures, one for inductive types, the other for linkages.

>>>>

We highlight some specialities of declarative style in QIIT that differ from a
conventional setting: 
(1) This declarative formulation of type theory does not
have an operational semantics but only judgemental equality (e.g. $\beta$-reduction is now $\beta$-equality).\footnote{Later we will prove
canonicity which can imply a certain level of computational ability of
our theory} 
(2) We directly use the equality from QIIT to represent judgemental equality in our syntax, relieve from us the burden to explicit mentioning equivalence closure; however, this also mean that, \textit{every piece of the syntax data}, including all kinds of judgements (QIIT type) and their derivations will have equality information upon them  
%We will recover canonicity afterwards, which implies the computational ability of our theory. 
% (3) We are dealing with quotient data, so we are actually manipulating
% \textit{equivalence classes of terms} instead of one representative (term) from the equivalence class.
(3) Instead of using meta-level substitution, we use explicit
substitution and de Bruijn indices in our intrinsic formulation. This is
also called \textit{substitution calculus} in the literature and is favoured due
to its closedness to categorical semantic and easir to handle in mechanization. We will go through these details one-by-one.

@@

% omitting presupposition
Due to our intrinsic setting, \textbf{we can and will omit a lot of premises of the typing rules}~(the presuppositions). 
For example, we don't explicitly
spell out that the rule \ruleref{Boolean} (that is stating a boolean is a proper type) requires a well-formed
context, even though it does implicitly inferrable, \textbf{because it is not
possible to have ill-formed(ill-typed) contexts in our intrinsic setting}. This is
a bit unusual for conventional extrinsic setting where they will have
well-formedness of context everywhere in most of the term judgements. We
consider this as one advantage of using intrinsic setting---this
``well-formed context'' condition is usually attached to some other premises thus
inferrable and can be considered as ``boilerplate code''. Free of
such boilerplate code, the intrinsic setting provides a more concise
representation.

@@

% Example : how judgemental equality is used for the universe
We review the concept of ``universe'' and at the same time showing how equality is representing semantic. In dependent type setting, function now can return type, of which the type is ``type of type'', i.e., \textit{the universe} $\cU$. We follow the style of \textit{type universe à la Tarski}~\cite{hofmann1997syntax} and
distinguish types from their \textit{codes or names} as showin in \ruleref{Code/Get Type}: given a \textit{code} (term) $\goodTerm{\Gamma}{T}{\cU}$ we can retrieve the corresponding type
$\goodType{\Gamma}{El~T}{}$, and vice-versa.

In other words, (1) a universe $\cU$ is a type including the \textit{codes or names} of types.
(2) a code (term) $\goodTerm{\Gamma}{T}{\cU}$ is not allowed to locate at type position
(i.e., at the rightmost position of a term judgement) but $El \ T$ is (i.e. we distinguish the code and the type). (3)
"c" is the ``inverse'' of $El$, obtaining  the \underline{"c"}ode of a
type. More concretely, using QIIT, the corresponding constructors will look like, (notice that the type on both sides of the equalities are always the same and thus a well-typed equality) 
\begin{align*}
  &\mintinline{Agda}{c : Ty Γ → Tm Γ U} &&\mintinline{Agda}{El : (a : Tm Γ U) → Ty } \\
  &\mintinline{Agda}{Elc : El (c T) ≡ T} && \mintinline{Agda}{cEl : c (El T) ≡ T}
\end{align*}
These two equations indicate how \textit{dynamic} should react when $El~"c"~T$ appears. Later we will see other similar $\beta$-equations but we won't display their (pseudo-)Agda formulation.

@@

$$
\quad 
\Rule[name=Func/DPair Subst]
{\goodSub{\Gamma}{\gamma}{\Delta}
\quad \goodType{\Delta}{A}{} 
\quad \goodType{\Delta, A}{B}{}
}
{
  \goodType{\Gamma}{(\Pi A B)[\gamma] \equiv \Pi A[\gamma] B[\gamma^\uparrow] }{j \cup k}
  \quad 
  \goodType{\Gamma}{(\Sigma A B)[\gamma] \equiv \Sigma A[\gamma] B[\gamma^\uparrow] }{j \cup k}
}
$$
%\YZ{What is the requirement on Delta?}
%\EDJreply{ Delta here is of type Con and thus Delta here should be a well-typed Context. This is another issue I want to talk about, in Agda these kinds of stuff is inferred (so the reader can infer them as well). For example, we know delta is at the place of context, so it is a context, but we don't need to specify that it is a ``well-typed context'' because every context is well-typed in this intrinsic-typing. What's more, I think adding this extra judgement here just cause significant code bloat (like what I complained to you before I know how to use implcit variable in agda. For example, TYPE SUBST, there will be two more judgements (with no more information there) ) (Another reason I prefer (pseudo) agda formulation) }\YZreply{I was thinking about 'Delta |- A'. I suppose it is also implicitly enforced via intrinsically typed syntax. Seems to me it is worth stating, however, or otherwise the rule appears to say that Delta can be any context as long as Gamma |- gamma : Delta.}\EDJreply{You are absolutely right! I think Delta |- A, even though inferrable, mentioning it is good, also some information are B is also ok, please check, resolve if ok, thanks!}
$$
\Rule[name=Type Subst]
{\goodType{\Delta}{T}{j} 
  \quad {\goodSub{\Gamma}{\gamma}{\Delta}}}
{\goodType{\Gamma}{T[\gamma]}{j}}
\quad
\Rule[name=Base Type Subst]
{\goodSub{\Gamma}{\gamma}{\Delta}}
{\goodType{\Gamma}{\cU[\gamma] \equiv \cU }{j + 1} \quad
  \goodType{\Gamma}{\cB[\gamma] \equiv \cB}{0} \quad 
  \goodType{\Gamma}{\bot[\gamma] \equiv \bot}{0}
}
$$
\judgebox{ \goodTerm{\Gamma}{t}{T} }
$$
\Rule[name=Code/Get Type]
{\goodType{\Gamma}{T}{j}}
{\goodTerm{\Gamma}{"c" \ T}{\cU}
}\quad
\Rule[]
{\goodTerm{\Gamma}{T}{\cU}}
{\goodType{\Gamma}{El \ T}{j}}
\quad 
\Rule[name=Term Subst]
{\goodTerm{\Delta}{t}{T}
  \quad {\goodSub{\Gamma}{\gamma}{\Delta}}}
{\goodTerm{\Gamma}{t[\gamma]}{T[\gamma]}}
$$
$$
\Rule[]
{\goodTerm{\Gamma, A}{t}{B}}
{\goodTerm{\Gamma}{\lambda t}{\Pi A B}}
\quad 
\Rule[]
{\goodTerm{\Gamma}{t}{\Pi A B}}
{\goodTerm{\Gamma, A}{"app"~t}{B}}
\quad 
\Rule[]
{\goodTerm{\Gamma}{u}{A} 
\quad \goodTerm{\Gamma}{v}{B[(id, u)]}}
{\goodTerm{\Gamma}{(u,v)}{\Sigma A B}}
\quad 
\Rule[]
{\goodTerm{\Gamma}{t}{\Sigma A B}}
{\goodTerm{\Gamma}{"pjl" \ t}{A}
\\\\ \goodTerm{\Gamma}{"pjr" \  t}{B[(id, "pjl" \  t)]}
}
$$
$$
\Rule
{}
{\goodTerm{\Gamma}{(\lambda t)[\gamma] \equiv \lambda (t[\gamma^\uparrow])}{\Pi A B}
\quad \goodTerm{\Gamma}{(u,v)[\gamma] \equiv (u[\gamma],v[\gamma])}{\Sigma A B}
\quad \goodType{\Gamma}{El \ (T[\gamma]) \equiv (El \ T) [\gamma]}{}
}
$$

$$
\Rule[name=Bool]
{}
{\goodTerm{\Gamma}{"tt", "ff"}{\cB}}
\quad
\Rule[]
{\goodSub{\Gamma}{\gamma}{\Delta}}
{\goodType{\Gamma}{(El \ T)[\gamma] \equiv El \ (T[\gamma]) }{j} \quad
 \goodTerm{\Gamma}{"tt"[\gamma] \equiv "tt"}{\cB} \quad 
 \goodTerm{\Gamma}{"ff"[\gamma] \equiv "ff"}{\cB} 
}
$$
\judgebox{\goodSub{\Gamma}{\sigma}{\Delta}}
$$
\Rule[name=Ept Subst]
{}{\goodSub{\Gamma}{\epsilon}{\cdot}}
\quad
\Rule[]
{}{\goodSub{\Gamma}{"id"}{\Gamma}}
\quad
\Rule[name=Sub Comp]{
  \goodSub{\Delta}{\delta}{\Theta}
  \quad \goodSub{\Gamma}{\gamma}{\Delta} 
}{\goodSub{\Gamma}{\delta \circ \gamma}{\Theta}}
\Rule[name=Sub Extend]
{\goodSub{\Gamma}{\sigma}{\Delta} \quad \goodTerm{\Gamma}{t}{A[\sigma]}}
{\goodSub{\Gamma}{(\sigma, t)}{(\Delta, A)}}
$$

\underline{\textit{\textbf{Substitution Calculus.}}}
We first review the substitution calculus.
Substitution is very (if not the most) important because (dependent) functions are
first-class citizens, and lambdas can wrap all possible terms and once we apply the functions,  \textbf{the substitution will happen on the wrapped body and the substitution needs to
commute} with all type/term constructs in the function body. However, our substitution \textbf{is part of the syntax} instead of a meta-operation in conventional setting. We review the key to substitution calculus, mainly surrounding the judgements $\goodSub{\Gamma}{\gamma}{\Delta},\goodTerm{\Gamma}{t}{T},\goodType{\Gamma}{T}{}$.

A term $\goodTerm{\Gamma}{t}{A}$ can be understood as one term $t$ of
type $A$, with a bunch of holes (of type $\Gamma$) inside the term $t$.
A \textbf{simultaneous substitution} $\goodSub{\Gamma}{\gamma}{\Delta}$ can be similarly
understood in the way that $\gamma$ will output a \textbf{series} of
(dependently-typed related) terms of types $\Delta$, parametrized by
$\Gamma$ (i.e., we can consider there are holes of type $\Gamma$ in $\gamma$).
Thus (1) \ruleref{Sub Extend} looks like concatenating one
series of terms with a new term (the weird looking type of
$\goodTerm{\Gamma}{t}{A[\sigma]}$ in the premise is due to our need of type $A$ in
the context $\Delta$ so that we can use \ruleref{Context Extension} to
create new context); (2) we have \ruleref{Proj Subst-1} to extract a
subpart of a given substitution; (3) \ruleref{Proj-Ext} explains why
substitution really looks like a (deeply-nested) pair.

@@

\underline{\textit{\textbf{Original MLTT Constructs.}}}
We only include  the dependent function types, dependent pair types, boolean type (as a base type later for canonicity), unit type, empty type and the universe type from the original MLTT. Dependent pair type is used to model Coq module, which is useful when we need to later model the ``compilation'' from
linkages to modules. We also have $\beta,\eta$-rule for both dependent
function type and dependent pair type.

For these two types using binder, de Bruijn indices emphasizes its difference from the conventional setting---we directly use $\lambda t$ as a function term where $t$ is the body, no binder indication is required---to refer to variables in the context, we use $\pi_1,\pi_2$ from substitution calculus instead of using their names in the conventional setting.  For example, the identity function $\lambda x. x$ is actually $\lambda(\pi_2 id)$. $"app"~t$ is even more different than that of the original setting---we don't require an application argument now. The reason is, (1) we asked for argument in conventional setting so we can use meta-level substitution as part of the dynamic, but here we use explicit substitution. If the user want to substitute some arguments, they can just use $[\gamma]$ instead. (2) Formulated in this way, "app" and $\lambda$ is symmetrical inverse to each other (just like how $El$ and "c" are to each other). This makes the $\beta\eta$-rule elegantly straightforward. Besides $\lambda$, $\Pi$ and $\Sigma$ are also using binders and are formulated in a similar style.

Apparently, all of the type constructs have to commute with substitution somehow. The substitution law for dependent function and dependent pair
requires \textit{de Bruijn lifting}: for a given substitution
$\goodTerm{\Gamma}{\sigma}{\Delta}$ and a type
$\goodType{\Delta}{A}{}$, we define its de Bruijn lifting
$\goodTerm{\Gamma,A[\sigma]}{\sigma^\uparrow := (\sigma \circ \pi_1, \pi_2)}{\Delta, A}$ . Intuitively speaking, de Bruijn lifting is to ``fill all the holes except the last one''.

\subsubsection{\textbf{Core Syntax of the Extended MLTT}}
\begin{figure}[!htb]
  \begin{align*}
&\sigma~:~"Sig"~\Gamma~n&&\mathboxtext{\goodSig{\Gamma}{\sigma}{n}}  &&\tau~:~"WSig"~\Gamma~n&&\mathboxtext{\goodWSig{\Gamma}{\tau}{n}} \\ &h~:~"Inh"~\Gamma~\sigma_1~\sigma_2&&\mathboxtext{\goodInh{\Gamma}{h}{\sigma}{\tau}} && f~:~"Seal"~\Gamma~\sigma~T&&\mathboxtext{\goodSeal{\Gamma}{f}{\sigma}{A}}\YZ{A little unfriendly to the reader that there are three different things that sigma can range over. Besides, later you also use gamma to range over substitutions and s to range over W-types, which add to the confusion.}\EDJreply{Fair point. I have fixed fix sigma and tau to be about signatures and W signature, and use gamma as substitutions}
  \end{align*}
\caption{Extended Judgement forms, as Agda types}
\end{figure}
\begin{figure}[!htb]
  \begin{minipage}[b]{0.3\linewidth}
      $$
      \Rule[name=Sig]
      {\goodCtx{\Gamma}{i} \quad n \in \nat}
      {\goodSig{\Gamma}{\_}{n} \text{ is well-formed}}
      $$
      $$
      \Rule[name=Seal]
      {\goodCtx{\Gamma}{i} \quad \goodSig{\Gamma}{\sigma}{n} 
      \quad \goodType{\Gamma}{A}{} }
      {\goodSeal{\Gamma}{\_}{\sigma}{A} \text{ is well-formed}}
      $$\YZ{Doesn't the left column show well-formedness rules for ONLY TWO of the four new judgments? Besides, are these rules used anywhere in the following? Seems a bit unusual to have well-formedness rules for judgments per se.}\EDJreply{Yes. Right column show all the well-formedness rule for all the judgement. That is unusual because it is a direct translation from QIIT. }
  \end{minipage}
  \begin{minipage}[b]{0.6\linewidth}
    \begin{minted}[]{agda}
      data Sig  : Con → ℕ → Set
      data WSig : Con → ℕ → Set 
      data Inh  : (Γ : Con) → Sig Γ n → Sig Γ m → Set
      Seal : (Γ : Con) → Sig Γ n → Sig Γ n → Set
    \end{minted}
  \end{minipage}

@@

The core language builds on MLTT and adds four new judgement forms.
% Among them, sealing judgement $\goodSeal{\Gamma}{\_}{\sigma}{A}$ is just a
% shorthand for term judgement
% $\goodTerm{\Gamma , \cC \sigma}{\_}{A[\pi_1]}$ where $A$ is a type.
% We have shown some exemplar rules for all four standard types of
% judgments. Now we focus on the newly introduced facility.

We have two kinds of signatures, one for inductive types, the other for linkages. We start with inductive type signature.

@@
